---
output: 
  html_document:
    keep_md: yes
    fig_caption: yes
---
#Reproducible research: Peer graded assignment 1

========================================================

**_1. Loading and preprocessing data_**

```{r, echo=TRUE}
dat <- read.csv("activity.csv")
dat$date <- as.Date(as.character(dat$date), format = "%Y-%m-%d")
```

**_2. Total Number of steps per day_**
```{r, echo=TRUE}
library(dplyr)
library(knitr)
sub1 <- subset(dat, steps >= 0, select = c(steps, date))
grouped <- group_by(sub1, date)
Totals <- summarise(grouped, Total.Steps = sum(steps, na.rm = TRUE))
kable(Totals, caption = "Total number of steps per day", align = "l", format = "markdown")
hist(Totals$Total.Steps, xlab = "Total steps", main = "Total number of steps", col = "blue", breaks = 8)
mean.med <- summarise(grouped, Mean = round(mean(steps, na.rm = TRUE)), Median = median(steps, na.rm = TRUE))
kable(mean.med, caption = "Mean and Median steps per day", align = "c", format = "markdown")
```

**_3. Average daily activity_**
```{r, echo=TRUE}
library(ggplot2)
sub2 <- subset(dat, steps >= 0, select = c(steps, interval))
grouped2 <- group_by(sub2, interval)
daily_interval_avg <- summarise(grouped2, Avg.Steps = mean(steps))
g <- ggplot(daily_interval_avg, aes(interval, Avg.Steps))
g+geom_line(colour = "purple")+ggtitle("Daily average steps taken per inteval")+ ylab("Average Steps")
i <- which.max(daily_interval_avg$Avg.Steps)
daily_interval_avg$interval[i]
```

The interval with maximun number of steps averaged by all days is `r daily_interval_avg$interval[i]`

**_4. Missing values_**
```{r, echo=TRUE}
icc <- !complete.cases(dat)
Nas <- sum(icc)
```
There are `r Nas` rows with missing values
```{r, echo=TRUE}
dat1 <- dat
m <- round(mean(dat1[,1], na.rm = TRUE))
for(i in 1: nrow(dat1)){
        if (is.na(dat1[i,1]) == TRUE){
            dat1[i,1] <- m
        }else{
                dat1[i,1] <- dat1[i,1]
        }
}
Totals1 <- summarise(group_by(subset(dat1, select = c(steps,date)),date), Total.Steps = sum(steps, na.rm = TRUE))
hist(Totals1$Total.Steps, xlab = "Total steps", main = "Total number of steps", col = "brown", breaks = 8)
mean.med1 <- summarise(group_by(subset(dat1, select = c(steps,date)),date), Mean = round(mean(steps, na.rm = TRUE)), Median = median(steps, na.rm = TRUE))
kable(mean.med1, caption = "Mean and Median steps per day", align = "c", format = "markdown")
```

**_5. Weekdays Vs Weekends_**

```{r, echo=TRUE}
dat1 <- mutate(dat1, Weekday = weekdays(dat1$date))
for(i in 1:nrow(dat1)){
        if(dat1[i,4] ==  "Saturday" | dat1[i,4] == "Sunday"){
                dat1[i,4] <- "weekend"
        }else if(dat1[i,4] == "Monday" | dat1[i,4] == "Tuesday" | dat1[i,4] == "Wednesday" | dat1[i,4] == "Thursday" | dat1[i,4] == "Friday"){
                dat1[i,4] <- "weekday"
        }
}
dat1$Weekday <- as.factor(dat1$Weekday)
```

```{r, echo= TRUE }
sub3 <- subset(dat1, select = c(steps, interval, Weekday))
grouped3 <- group_by(sub3, Weekday, interval)
interval_Avg.Steps <- summarise(grouped3, Avg.Steps = mean(steps))
g <- ggplot(interval_Avg.Steps, aes(interval, Avg.Steps))
g <- g + geom_line(colour = "brown") +facet_grid(Weekday ~.)
g <- g+ xlab("Interval") + ylab("Number of Steps") + ggtitle("5-minute interval average number of steps taken")
g
```

* Weekends have more steps taken across most 5-minute intervals than weekdays
* Weekdays have the interval with the highest number of steps 